---
title: "Course Project 1"
author: "cylevass"
date: "29 octobre 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(dplyr)
```

## Loading and preprocessing the data
```{r}
my_data <- read.csv("activity.csv")
my_data$hour <- substr(sprintf("%04d",my_data$interval),1,2)
my_data$minute <- substr(sprintf("%04d",my_data$interval),3,4)
my_data$mytime <- as.POSIXct(paste(my_data$date,my_data$hour,my_data$minute,sep="-"),"%Y-%m-%d-%H-%M",tz="")
head(my_data)
```

## Total number of steps taken per day
```{r}
g <- ggplot(data=my_data, aes(my_data$mytime, weight = my_data$steps)) +
    geom_histogram(bins=61, fill="blue", col="white", alpha=0.7) +
    scale_x_datetime(breaks = date_breaks("1 week"), date_labels = "%Y-%m-%d") +
    xlab("Date") +
    ylab("Steps per day") +
    ggtitle("Total of steps per day") +
    theme(axis.text.x = element_text(size=8, angle=45, hjust=1))
g
tot_per_day <- tapply(X = my_data$steps, INDEX = my_data$date, FUN = sum)
```

Mean number of steps per day :
```{r}
mean(tot_per_day, na.rm=TRUE)
```
Median number of steps per day :
```{r}
median(tot_per_day, na.rm=TRUE)
```

## Average daily activity pattern
```{r}
mean_per_interval <- tapply(X = my_data$steps, INDEX = paste(my_data$hour,my_data$minute,sep=':'), FUN = mean, na.rm=TRUE)
plot(mean_per_interval,
     xaxt = "n",
     type = "l",
     main = "Mean number of steps per 5-minutes",
     xlab = "5-minutes interval",
     ylab = "Mean number of steps")
axis(1, at=1:length(mean_per_interval), labels=names(mean_per_interval))
```

Maximum number of steps on average :
```{r}
names(which.max(mean_per_interval))
max(mean_per_interval)
```


## Missing values

Number of missing values :
```{r}
sum(is.na(my_data$steps))
```

Filling the missing values with the average of steps for the intervals that miss :
```{r}
my_data_2 <- my_data
my_data_2 %>%
 group_by(interval) %>%
 mutate(steps = (ifelse(is.na(steps), as.numeric(mean(steps, na.rm = TRUE)), steps))) -> my_data_2
head(my_data_2)
```

New plot :
```{r}
g <- ggplot(data=my_data_2, aes(my_data_2$mytime, weight = my_data_2$steps)) +
    geom_histogram(bins=61, fill="blue", col="white", alpha=0.7) +
    scale_x_datetime(breaks = date_breaks("1 week"), date_labels = "%Y-%m-%d") +
    xlab("Date") +
    ylab("Steps per day") +
    ggtitle("Total of steps per day") +
    theme(axis.text.x = element_text(size=8, angle=45, hjust=1))
g
tot_per_day <- tapply(X = my_data_2$steps, INDEX = my_data_2$date, FUN = sum)
```

Mean number of steps per day :
```{r}
mean(tot_per_day, na.rm=TRUE)
```
Median number of steps per day :
```{r}
median(tot_per_day, na.rm=TRUE)
```

==> impact of the missing values is neglectable in this case.

## Differences in activity patterns between weekdays and weekends

```{r}
my_data_3 <- my_data
my_data_3$day_type <- weekdays(as.Date(my_data_3$date))
my_data_3$day_type[my_data_3$day_type=="lundi"] <- "weekday"
my_data_3$day_type[my_data_3$day_type=="mardi"] <- "weekday"
my_data_3$day_type[my_data_3$day_type=="mercredi"] <- "weekday"
my_data_3$day_type[my_data_3$day_type=="jeudi"] <- "weekday"
my_data_3$day_type[my_data_3$day_type=="vendredi"] <- "weekday"
my_data_3$day_type[my_data_3$day_type=="samedi"] <- "weekend"
my_data_3$day_type[my_data_3$day_type=="dimanche"] <- "weekend"
my_data_3$day_type <- factor(my_data_3$day_type)

mean_per_interval_day <- tapply(X = my_data_3$steps, INDEX = paste(my_data_3$day_type,my_data$hour,my_data$minute,sep=':'), FUN = mean, na.rm=TRUE)

plot(mean_per_interval_day,
     xaxt = "n",
     type = "l",
     main = "Mean number of steps per 5-minutes",
     xlab = "5-minutes interval",
     ylab = "Mean number of steps")
axis(1, at=1:length(mean_per_interval_day), labels=names(mean_per_interval_day))

```

